---
title: "Global Biodiversity, Trade, and Conservation Funding"
author: "Evan Daisy, Sanjana Sunder, Kim Zhou"
date: "05/13/2021"
output:
  rmdformats::readthedown:
    thumbnails: false
    highlight: NULL
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.align="center")
library(tidyverse)
library(datasets)
library(viridis)
library(maps)
library(leaflet)
library(ggnetwork)
library(igraph)
library(bslib)
library(rgdal)
library(sqldf)
library(formattable)
library(xtable)
library(kableExtra)
```

## Introduction
Over the past 40 years, vertebrate population sizes measured by the World Wildlife Fund have more than halved, mostly due to overhunting and habitat destruction from urban development, and research out of the University of Arizona suggests that one in three species are likely to be extinct by the end of the century (article [here](https://www.cnn.com/2020/02/21/weather/species-extinction-climate-trnd/index.html)), in large part due to human activities. Our motivation for this blog post is to analyze global species biodiversity distribution and our impact on the loss and preservation of said biodiversity. Maintaining natural ecosystem biodiversity is crucial to sustaining human life on this planet because functioning ecological systems provide clean air and water, as well as pollination of agriculture crops. Without the natural species present, the delicate balance of these systems is destroyed, and with it, our planet and us.

The first section of this blog post is dedicated to analyzing current global vertebrate species diversity using data from the World Wildlife Fund Living Planet Index (see below for all sources). By analyzing the countries based on the number of species that fall in IUCN threat categories (such as critically endangered, least concern, etc.), we gain an understanding of where we should be directing conservation funding and efforts because it is the ecosystems in these areas that are most in need of saving and are the most susceptible to loss. Here we note that this information is not the most comprehensive since the data is solely focused on vertebrate species, while most of the world’s overall biodiversity includes invertebrate species. We also compare the distribution of species to the global distribution of conservation funding using data from published data on tracking conservation funding on the Dryad website. By comparing the two maps, we can determine which countries need more funding directed towards conservation and which countries already dedicate significant amounts of money towards conservation efforts.

Next, we attempt to quantify an aspect of our impact on biodiversity, which takes the form of international animal trade. Using data from the Convention on International Trade in Endangered Species of Wild Fauna and Flora, or CITES for short, we analyze the legal trade of animals and animal products globally to determine which countries were playing the largest role in the depletion of animal species. Unfortunately, we were limited to legal trade, so it is important to note that much of species loss, especially of endangered species, comes from illegal trade. 

We finally shift our focus to the US and look at both species diversity in national parks throughout the country and at major cities in the East using data from the National Parks Service and a published paper, respectively. The trends in major city species diversity illustrate the impact of urbanization on the introduction of non-native and loss of native species.

## World Maps

For interactive versions of these maps, click [here](https://edaisy21.shinyapps.io/blog_app/)

### Biodiversity

```{r, echo = FALSE}
threatened_iucn <- read.csv("threatenediucn_final.csv")
world_map <- map_data(map = "world"
                      , region = ".")

threatened_all <- threatened_iucn %>%
  group_by(Country) %>%
  summarise(tot_species = sum(num_species)) %>%
  full_join(world_map, by = c("Country" = "region"))

ggplot(data = threatened_all, aes(x = long, y = lat, group = group)) +
      geom_polygon(mapping = aes(x = long, y = lat, fill = tot_species), 
                   colour = "black") +
      theme_void()  +
      coord_fixed(ratio = 1.3) +
      scale_fill_viridis(option = "magma", direction = -1, na.value = "white") +
      labs(title = "Global Species Diversity"
           , fill = "Number of Species*"
           , caption = "*Regions in white have no data")
```

This figure depicts the total number of species by country that are included in the IUCN Redlist. Note that this clearly is not an accurate representation of global biodiversity. Tropical regions are known to harbor far greater biodiversity than more temperate regions, such as most of Europe or the United States. Thus the fact that the United States and Canada are displayed as having the greatest species richness here, followed by places like Australia, could be a function of the area of these countries, but that alone would not be sufficient to explain the disparity, because places like Russia and China do not display the same unusually high species richness. This is probably due then to the amount of available resources that each country devotes to the tabulation of species data, because tropical regions may not have as many resources and researchers as the United States to devote to the counting of species, and places like China and Russia may not prioritize it as heavily. Additionally, it appears that there is not much distinction in biodiversity of the rest of the world compared to the United States and Canada because the color scale is relative and the number of recorded species in these countries is so much greater that it skews the coloring. Therefore, leaving Canada and the United States out of the calculations would give a more comprehensive depiction of the rest of the world's recorded species diversity. 

### Conservation Funding (as of 2013)

```{r, echo = FALSE, message = FALSE, warning = FALSE, class.source = "greenCode"}
conservation <- read_csv("conservation_final.csv")

# Map of Log(funding)
ggplot(conservation, aes(x = long, y = lat, group = group, 
                         fill = log_funding)) +
  geom_polygon(color = "black") +
  theme_void() +
  coord_fixed(ratio = 1.3) +
  labs(title = "Conservation Funding By Country*"
       , subtitle = "as of 2013"
       , fill = ""
       , caption = "*Regions in white have no data") +
  #scale_fill_continuous(low="thistle2", high="darkred",guide="colorbar", na.value="white")
  scale_fill_viridis(option = "viridis", direction = -1, na.value="white")
```

This figure depicts the amount of funding each country is granted and/or devotes to conservation, on a log scale. The countries that receive and utilize the most conservation funding are the same countries that are displayed above as having the greatest biodiversity. This is not surprising, in that it is probably since both of these variables (conservation funding and observed species counts) are a function of the amount to which a country is able to and chooses to prioritize conservation efforts. We know that these countries are not in fact the most biodiverse however, so although it is good that they prioritize conservation within their borders, the fact that conservation in Africa and Central America is underfunded in comparison should be a cause for concern. Additionally, although it appears countries like the United States are doing a great job allocating funding for conservation, this map is relative, so it is not to say we shouldn't be dedicating more funding to the preservation of habitats and species. 

```{r, echo = FALSE}
ggplot(conservation, aes(x = long, y = lat, group = group, 
                         fill = log_fpgdp)) +
  geom_polygon(color = "black") +
  theme_void() +
  coord_fixed(ratio = 1.3) +
  labs(title = "Log(Conservation Funding as a Fraction of total GDP) by Country*"
       , subtitle = "as of 2013"
       , fill = ""
       , caption = "*Regions in white are missing GDP or funding data") +
  #scale_fill_continuous(low="thistle2", high="darkred",                      guide="colorbar", na.value="white")
  scale_fill_viridis(option = "viridis", direction = -1, na.value="white")
```

This figure shows that the disparity between the US, Canada, and other countries around the world is not so large when we factor in GDP. It shows that countries like the Central African Republic are in fact receiving and using funds for conservation that are significant relative to their GDP's. When we look at the funding in this light, conservation in Asian countries seems to be underfunded relative to their GDP's, despite the biodiversity of the continent.

### Network of Trade between countries

```{r, echo = FALSE}
 # Render the Network of trade between countries
    data <- read.csv("./trade.csv")
    trade_coun <- data %>%
                  select(importer_country, exporter_country) %>%
                  group_by(importer_country, exporter_country) %>%
                  summarize(trade_occurrence = n(), .groups = 'drop') %>%
                  filter(!(is.na(importer_country) | importer_country == "") &
                           !(is.na(exporter_country) | exporter_country == "") ) %>%
                  filter(trade_occurrence >= 200) %>%
                  arrange(desc(trade_occurrence))
    
    net <- graph_from_data_frame(trade_coun, directed = TRUE)
    net_plot <- ggnetwork(net)
    ggplot(data = net_plot
           , aes(x = x, y = y, xend = xend, yend = yend)) +
      geom_edges(curvature = 0.1, arrow=arrow(type="closed", length=unit(8,"pt"))
                 , aes(color = trade_occurrence)) +
      geom_nodes() +
      geom_nodelabel(aes(label = name)) +
      theme_blank() +
      labs(color = "Total Occurrence") +
      scale_color_continuous(type = "viridis")
```

This figure depicts the trade of species between countries (both imported and exported) and the count of species traded between two countries as a network plot.  The color of the edges between the countries indicate the frequency, as trade between countries numerous, the trade occurrence of over 200 species between the countries is represented.  

### Count of Species exported by the countries

```{r, , echo = FALSE}
# data vector to extract the total species exported by each country
  exp_country <- data %>%
      select(exporter_country, exporter) %>%
      group_by(exporter, exporter_country) %>%
      summarize(count = n(), .groups = 'drop') %>%
      filter(!(is.na(exporter_country) | exporter_country == "")) %>%
      filter(str_detect(exporter_country, "[a-z]")) 
    
    # Read this shape file with the rgdal library. 
    world_spdf <- readOGR( 
      dsn = paste0(getwd(),"") , 
      layer = "TM_WORLD_BORDERS_SIMPL-0.3",
      verbose = FALSE
    ) 
    # join data frame with shape file
    data_chloro <- world_spdf@data %>%
                    full_join(exp_country, by = c("ISO2" = "exporter"))

    mypalette <- colorBin( palette = "YlOrBr", domain = data_chloro$count
                           , na.color = "transparent", bins = 6)
    mytext <- paste(
      "Country: ", data_chloro$exporter_country,"<br/>", 
      "Count of Species Exported: ", data_chloro$count, "<br/>", 
      "Category: ", "Appendix I", 
      sep = "") %>%
      lapply(htmltools::HTML)
    
    # Basic choropleth with leaflet?
   plot <- leaflet(world_spdf) %>% 
      addTiles()  %>% 
      setView( lat=10, lng=0 , zoom=2) %>%
      addLegend( pal = mypalette, values = ~data_chloro$count
                 , opacity=0.9, title = "Count"
                 , position = "bottomleft" ) %>%
      addPolygons(
        fillColor = ~mypalette(data_chloro$count), 
        layerId = ~data_chloro$exporter_country,
        stroke = FALSE, 
        smoothFactor = 0.5,
        fillOpacity = 0.9, 
        color = "white", 
        weight = 0.3,
        label = mytext,
        labelOptions = labelOptions( 
          style = list("font-weight" = "normal", padding = "3px 8px"), 
          textsize = "13px", 
          direction = "auto"))
   plot
```
This figure represents a chloropleth map that indicates the total species exported by each country. When a specific country is chosen, the taxonomy of the species exported is represented by a table as well as a bar plot gives the count of species exported categorized by Appendix. An interactive version of the plot click [here]( https://sanj-sund.shinyapps.io/Blog-Team-Save-the-Planet-main/)


## United States Data

One takeaway from the maps above is that the United States has the funding and the drive to record large quantities of data about species diversity. Below are some charts that allow us to see differences in biodiversity between places within the United States.
  
### National Parks

```{r, echo = FALSE, message = FALSE}
np_species_by_area <- read_csv("np_tot_species.csv")

ggplot(data = np_species_by_area, aes(x = reorder(park_name, n), 
                                      y = n)) +
                geom_bar(stat = "Identity", aes(fill = park_name)) + 
                theme(axis.text.x = element_text(angle = 90, 
                                                 hjust = 1)) +
                theme(legend.position = "none") +
                labs(title = "Total Number of Species By Park") +
                xlab("Park") + 
                ylab("Total Species")
```

### Northeastern Cities

```{r, echo = FALSE, message = FALSE}
cities <- read_csv("cities_final.csv")
ggplot(cities, aes(x = City, y = Total)) + 
  geom_bar(stat = "Identity", aes(fill = City)) +
  labs(title = "Total Species By City") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  ylab("Total Species")
```

```{r, echo = FALSE}
ggplot(cities, aes(x = City, y = Species)) + 
  geom_bar(stat = "Identity", aes(fill = Origin)) +
  labs(title = "Native and Introduced Species By City") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  ylab("Total Species")
```

It makes sense that more species have been introduced to the four coastal cities than to the four inland cities, considering how many invasive species are transported (intentionally or not) by boats. For future studies, it would be worthwhile to continue these comparisons across more Midwest cities as well as cities along the West coast.

## Conclusion
The largest problem we encountered was finding data for the questions we wanted to address. There was no readily available access to species diversity of invertebrates, most likely because of the sheer number of these species. However, future work can be done to record this data and reconstruct our global distribution for a more accurate understanding of global biodiversity. Additionally, we could not find published data on illegal animal trade, and it is difficult to imagine this data even exists, which limits our analysis of human trade impact on species. 

Furthermore, to address some of the questions asked in class, while we did not account for GDP of the country in plotting available diversity data, this is a great way to expand on our current work because taking GDP into consideration can better standardize the number of species by accounting for resources a country can dedicate to record species diversity. Hence, we more accurately see if the US, for example, actually has significantly more species, or that it has more resources to properly track the number of species. 

## Data Sources
* WWF biodiversity (downloadable [csv](https://www.dropbox.com/s/t48dd7m0o088w69/LPI_pops_restricted_20160421.csv?dl=0))

* CITES trade (downloadable [csv](https://www.kaggle.com/cites/cites-wildlife-trade-database)) 


* Conservation funding (downloadable [csv](https://datadryad.org/stash/dataset/doi:10.5061/dryad.p69t1)


* National Parks (downloadable [csv](https://www.kaggle.com/gsdeepakkumar/biodiversity-an-eda/data?select=species.csvgbif.org%2Fdataset%2F67c54f85-7910-4cbf-8de4-6f0b136a0e34)) 

* Urban species (downloadable [excel](http://www.urbanhabitats.org/v01n01/speciesdiversity_full.html))
